---
title: "Diamonds"
author: "Amjad Darwish"
date: "3/15/2017"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE)
```

##Introduction
This document will examine the nature of diamond **pricing** when wieghed against things like **Carats**, **Color**, **Cut**, and **Clarity**.

## Price Relative to Carats
```{r , echo=FALSE, message=FALSE}
ggplot(data = diamonds) + geom_point(aes(x= price, y= carat)) + geom_smooth(aes(x=price, y=carat))
```

Looking at the data this way will be very difficult for us due to the huge data points it includes. We need to filter and play around with the dataset to be able to come to any meaningful conclusions.

## Common Colors
Let's first start off with colors. We would want to focus on the most used or common diamond colors. Below is a graph that shows the number of diamonds presented by each color.
```{r , echo=FALSE, message=FALSE}
ggplot(data = diamonds) +
  geom_bar( aes(color, fill = color)) + scale_fill_brewer(palette = "YlOrRd")
```

As shown in the above graph, the most common diamond colors in the market are *E*, *F*, and *G*. Let's try and explore more information regarding those specific colors.

```{r, include=FALSE}
diamonds2 <- diamonds %>% filter(color %in% c("E", "F", "G"))
```

## Premium Cuts
Diving deeper into the three colors we talked about earlier, we can conclude from the graph below that the most common cut for all of them is the ideal cut, which ranks as the **highest quality** cut there is.

```{r, echo=FALSE}
ggplot(data = diamonds2, aes(x = "cut share", fill = cut)) +
  geom_bar(width = 1) + coord_polar("y", start=0) + scale_fill_brewer(palette="Blues") +
  theme_minimal()
```


Further asserting that notion is the fact that, as mentioned below, these three colors are cut mostly in *Ideal*.

```{r, echo=FALSE}
ggplot(data = diamonds2) +
  geom_bar(aes(cut), fill = "steelblue2") + facet_grid(color ~ .)
```

## Clarity Levels

```{r include=FALSE}
diamonds3 <- diamonds2 %>% filter(cut %in% c("Ideal", "Premium", "Very Good"))
```

```{r echo=FALSE, message=FALSE}
ggplot(diamonds3, aes(price, carat)) +
  geom_point(aes(colour = clarity)) + geom_smooth(aes(x=price, y=carat), color = "gray37") + scale_colour_brewer(palette = "PRGn")
```




